{ \fontfamily{ptm}\selectfont First Order Nonsmooth Optimization: Algorithm Design, Variational analysis, and Applications }

### 2024-12-30 Notebook LaTEX/Proposal/main.tex
>w<, changes

\section{Introduction}
    Let $\RR^n$ be the ambient space. We consider 
    \begin{align}
        \min_{x \in \RR^n} \left\lbrace
            F(x): f(x) + g(x)
        \right\rbrace.
    \end{align}\label{eqn:additive-comp-obj}
    Unless specified, assume $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex and $g:Q \rightarrow \overline \RR$ is convex. 
    This type of problem is referred to as additive composite problems in the literature. 
    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had two lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithms?
        \item Is it possible to attain faster convergence rate without knowledge about the strong convexity of function $f$?
        \item Is it possible to describe the convergence of function value for momentum sequences that are much weaker than the Nesterov's rule? 
    \end{enumerate}
    The good news is we have definitive answers for all questions by our own efforts of research. 
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:spectral-momentum} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:spectral-momentum} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par
    Section \ref{sec:catalyst} are results of literatures review in MATH 590. 
    \todo{$\square$ Add Lin's papers and Paquette's papers.  }{\hl{It's based on a series of papers}} in the topic of Catalyst Meta Acceleration method for First Order Variance Reduced Methods. 
    We will point out potential future direction of research of Catalyst acceleration. 
    \par
    Section \ref{sec:pep}, \ref{sec:inexact-prox}, \ref{sec:nes-acc-ncnvx} preview literatures in  nonsmooth optimization frontier research where progress and impacts can be made.  
    \subsection{Theme of the research}
        This section specifies a theme of the research in this proposal. 
        Out first objective is to explore the Goldilocks zones between these topics: theories of variational analysis, design of continuous optimization algorithm and applications in sciences, engineering, and statistics.  
        Our second objective is to identify the ``chemistry'' occurring between properties of functions and the designs of continuous optimizations algorithm and how it impacts the convergence and behaviors of the algorithms.

@@

@@

        \todo{$\square$ Add Guler's 1992 new proximal point paper. \\ $\square$ Cite it. }{\hl{Güler}} 

@@

        \todo{$\square$ Add the paper: ``Accelerating the cubic regularization of Newton’s method on convex
        problems''\\ $\square$ Cite it.}{\hl{Nesterov}}
        also used the techinique to design an accelerated cubic regularized Newton's method. 

@@

        \todo{$\square$ Cite it. }\hl{Chapter 6 of Nesterov's book}, it's also used to derive a method of accelerated mirror descent. 

@@

        \todo{$\square$ Add paper: ``Accelerated regularized newton methods for
        minimizing composite convex functions'' \\ $\square$ Cite it. }{\hl{Geovani N. et al}}

@@

    The proofs haven't been carefully verified by authoratative people other than the author yet. 

@@

    \todo{$\square$ Add Nesterov's original paper. \\ $\square$ Cite it.  }{\hl{Proposed back in 1983 the original Nesterov's acceleration method}} 

@@

    \todo{$\square$ Add Nesterov's new book. \\ $\square$ Cite it.}{\hl{This convergence rate is considered optimal for all class of $L$ Lipschitz smooth convex function}}. 

@@

    \todo{$\square$ Add Beck 2009 FISTA original paper. $\square$ Cite it.}{\hl{Beck and Toubelle}} 

@@

    \todo{$\square$ Add Necoara linear convergence \\ $\square$ Add Aujol 2024 Parameter free FISTA restart. \\ $\square$ Cite Section 5.2.2 for the former, cite the entirety for the later. }{\hl{See}} 

@@

    \todo{$\square$ Add Neol J. Walkington's ``Nesterov’s Method for Convex  Optimization''. \\ $\square$ Cite it. }{\hl{Walkington's writing}}, 

@@

    \todo{$\square$ Add Beck's first order textbook\\ $\square$ Cite 10.7.7. \\ $\square$ Add Nesterov's textbook. \\ $\square$ Cite it. }{\hl{The same variant also appears in Beck's book as V-FISTA, and Nesterov's book as (2.2.22).}} 

@@

    \todo{$\square$ Add paper: ``On the convergence of the iterates of the...''\\ $\square$ Cite them. }{\hl{Chambolle, Dossal}} 

@@

        \todo{$\square$ Add ``On the convergence of the iterates of...'' \\ $\square$ Cite them. }{\hl{Chambolle, Dossal}}, 

@@

        \todo{$\square$ Add beck's original FISTA Paper. \\$\square$ Cite it. }

@@

        Organization now follows. 
        Section \ref{ssec:building-block-rwapg} provides a stepwise description of the R-WAPG iterative algorithm along with an inequality crucial to proving the convergence rate later. 
        Section \ref{ssec:rwpag-seq-and-alg} introduce the definition of an R-WAPG sequence, which constraints all possible parameters used permitted by the algorithm.
        The section also states the full R-WAPG algorithm and an upper bound on $F(x_k) - F(x^*)$. 
        Here, $x_k$ is generated by the R-WAPG algorithm and $x^*$ is the minimizer. 
        Section \ref{ssec:quiv-repr-rwapg} bring forward three equivalent forms of the R-WAPG algorithm making it more comparable with other Acceerated Proximal Gradient method appeared in the literatures. 
        Section \ref{ssec:describe-variants-with-rwapg} gives characterizations of specific R-WAPG sequence that leads to convergence of the R-WAPG algorithm in terms of the optimality gap. 
        The section also identifies specific instance of permissible R-WAPG sequences where it fits with the FISTA, V-FISTA, and the algorithm proposed by 
        \todo{$\square$ Cite the Chambolle, Dossal 2015 paper here again. }{\hl{Chambolle Dossal}}.

@@

    \subsection{Equivalent forms of R-WAPG algorithm}\label{ssec:quiv-repr-rwapg}
        \todoinline{
            Definitions:
            \begin{enumerate}
                \item[$\blacksquare$] R-WAPG Intermediate form.
                \item[$\blacksquare$] R-WAPG Similar triangle form. 
                \item[$\blacksquare$] R-WAPG Momentum form. 
            \end{enumerate}
        }
        \todoinline{
            Theorems: 
            \begin{enumerate}
                \item[$\blacksquare$] R-WAPG First equivalent form. 
                \item[$\blacksquare$] R-WAPG Second equivalent form.
                \item[$\blacksquare$] R-WAPG Third equivalent form.
            \end{enumerate}
            Lemmas
            Todo: 
            \begin{enumerate}
                \item Explain what is what. 
            \end{enumerate}
        }
        Definitions \ref{def:r-wapg-intermediate}, \ref{def:r-wapg-st-form} and \ref{def:r-wapg-momentum-form} are three equivalent representations of the R-WAPG algorithms. 
        Propositions \ref{prop:wapg-first-equivalent-repr}, \ref{prop:wagp-st-form} and \ref{prop:r-wapg-momentum-repr} states the equivalences between the forms and the sufficient conditions for initial conditions of $x_1, v_1$ such equivalence holds. 
        \todo{$\square$ Add and explain this part. }{\hl{
        The remarks of the definitions identifies specific instances in the literatures where the Accelerated Proximal Gradient method were presented under this specific form. 
        }}

@@

@@

@@

>>>>

\section{Introduction}
    Let $\RR^n$ be the ambient space. We consider 
    \begin{align}
        \min_{x \in \RR^n} \left\lbrace
            F(x): f(x) + g(x)
        \right\rbrace.
    \end{align}\label{eqn:additive-comp-obj}
    Unless specified, assume $f:\RR^n \rightarrow \RR$ is $L$-Lipscthiz smooth $\mu \ge 0$ strongly convex and $g:Q \rightarrow \overline \RR$ is convex. 
    This type of problem is referred to as additive composite problems in the literature. 
    \par
    Our ongoing research concerns accelerated proximal gradient type method for solving (\ref{eqn:additive-comp-obj}). 
    In the expository writing by Walkington \cite{noel_nesterovs_nodate}, a variant for of accelerated gradient method for strongly convex function $f$ is discussed. 
    We had two lingering questions after reading it. 
    \begin{enumerate}
        \item Do there exist a unified description for the convergence for both variants of the algorithms?
        \item Is it possible to attain faster convergence rate without knowledge about the strong convexity of function $f$?
        \item Is it possible to describe the convergence of function value for momentum sequences that are much weaker than the Nesterov's rule? 
    \end{enumerate}
    The good news is we have definitive answers for all questions by our own efforts of research. 
    Section \ref{sec:unify-nes-acceleration}, \ref{sec:spectral-momentum} are our ongoing research which present the answers to the questions. 
    \par
    In Section \ref{sec:unify-nes-acceleration}, we proposed the method of ``Relaxed Weak Accelerated Proximal Gradient (R-WAPG)'' as the foundation to describe several variants of Accelerated proximal gradient method in the literatures. 
    The convergence theories of R-WAPG allows us to model convergence of accelerated proximal gradient method where the momentum sequence doesn't strictly follow the conditions presented in the literatures. 
    The descriptive power of R-WAPG allows convergence analysis for all the variants using one single theorem. 
    \par
    In Section \ref{sec:spectral-momentum} we propose a practical algorithm that exploits a specific term in the proof of R-WAPG to achieve faster convergence for solving (\ref{eqn:additive-comp-obj}) without knowing parameter $L, \mu$ in prior. 
    Results of numerical experiments are presented. 
    \par
    Section \ref{sec:catalyst} are results of literatures review in MATH 590. 
    \todo{$\blacksquare$ Add Lin's papers and Paquette's papers.  }{\hl{It's based on a series of papers}} in the topic of Catalyst Meta Acceleration method for First Order Variance Reduced Methods. 
    We will point out potential future direction of research of Catalyst acceleration. 
    \par
    Section \ref{sec:pep}, \ref{sec:inexact-prox}, \ref{sec:nes-acc-ncnvx} preview literatures in  nonsmooth optimization frontier research where progress and impacts can be made.  
    \subsection{Theme of the research}
        This section specifies a theme of the research in this proposal. 
        Out first objective is to explore the Goldilocks zones between these topics: theories of variational analysis, design of continuous optimization algorithm and applications in sciences, engineering, and statistics.  
        Our second objective is to identify the ``chemistry'' occurring between properties of functions and the designs of continuous optimizations algorithm and how it impacts the convergence and behaviors of the algorithms.

@@

            Definition taken from Definition 4.3.1 from 
            \todo{$\blacksquare$ Add bib \\ $\square$ Cite. }\hl{Pang, Cui's book}

@@

        \todo{$\blacksquare$ Add Guler's 1992 new proximal point paper. \\ $\square$ Cite it. }
        {\hl{Güler}} 

@@

        \todo{$\blacksquare$ Add the paper: ``Accelerating the cubic regularization of Newton’s method on convex
        problems''\\ $\square$ Cite it.}
        {\hl{Nesterov}}
        also used the technique to design an accelerated cubic regularized Newton's method. 

@@

        \todo{$\blacksquare$ Cite it. }\hl{(6.1.19) of Nesterov's book }\cite{nesterov_lectures_2018}, it's also used to derive a method of accelerated mirror descent. 

@@

        \todo{$\blacksquare$ Add paper: ``Accelerated regularized newton methods for minimizing composite convex functions'' \\ $\square$ Cite it. }
        {\hl{Geovani N. et al}}

@@

    The proofs haven't been carefully verified by authoritative people other than the author yet. 

@@

    \todo{$\blacksquare$ Add Nesterov's original paper. \\ $\square$ Cite it.  }
    {\hl{Proposed back in 1983 the original Nesterov's acceleration method}} 

@@

    \todo{$\blacksquare$ Add Nesterov's new book. \\ $\blacksquare$ Cite it.}
    {\hl{This convergence rate is considered optimal for all class of $L$ Lipschitz smooth convex function }}\cite{nesterov_lectures_2018}. 

@@

    \todo{$\blacksquare$ Add Beck 2009 FISTA original paper. $\square$ Cite it.}
    {\hl{Beck and Toubelle}} 

@@

    \todo{
        $\blacksquare$ Add Necoara linear convergence 
        $\blacksquare$ Add Aujol 2024 Parameter free FISTA restart. 
        $\square$ Cite Section 5.2.2 for the former, cite the entirety for the later. 
    {\hl{See} (5.2.2) Necoara et al. and Aujol et al. } 

@@

    \todo{
        $\blacksquare$ Add Neol J. Walkington's ``Nesterov's Method for Convex  Optimization''. \\ $\square$ Cite it. 
    {\hl{Walkington's writing}}, 

@@

    \todo{
        $\blacksquare$ Add Beck's first order textbook
        \\ $\square$ Cite 10.7.7. 
        \\ $\blacksquare$ Add Nesterov's textbook. 
        \\ $\square$ Cite it. }
    {\hl{
        The same variant also appears in Beck's book as V-FISTA, and Nesterov's book as (2.2.22).
    }} 

@@

    \todo{
        $\blacksquare$ Add paper: ``On the convergence of the iterates of the...''
        \\ $\square$ Cite them. }
    {\hl{Chambolle, Dossal}} 

@@

        \todo{
            $\blacksquare$ Add ``On the convergence of the iterates of...'' \\ $\square$ Cite them.}
        {\hl{Chambolle, Dossal}}, 

@@

        \todo{
            $\blacksquare$ Add beck's original FISTA Paper. \\$\square$ Cite it.

@@

        Organization now follows. 
        Section \ref{ssec:building-block-rwapg} provides a stepwise description of the R-WAPG iterative algorithm along with an inequality crucial to proving the convergence rate later. 
        Section \ref{ssec:rwpag-seq-and-alg} introduce the definition of an R-WAPG sequence, which constraints all possible parameters used permitted by the algorithm.
        The section also states the full R-WAPG algorithm and an upper bound on $F(x_k) - F(x^*)$. 
        Here, $x_k$ is generated by the R-WAPG algorithm and $x^*$ is the minimizer. 
        Section \ref{ssec:quiv-repr-rwapg} bring forward three equivalent forms of the R-WAPG algorithm making it more comparable with other Accelerated Proximal Gradient method appeared in the literatures. 
        Section \ref{ssec:describe-variants-with-rwapg} gives characterizations of specific R-WAPG sequence that leads to convergence of the R-WAPG algorithm in terms of the optimality gap. 
        The section also identifies specific instance of permissible R-WAPG sequences where it fits with the FISTA, V-FISTA, and the algorithm proposed by 
        \todo{$\square$ Cite the Chambolle, Dossal 2015 paper here again. }
        {\hl{Chambolle Dossal}}.

@@

    \subsection{Equivalent forms of R-WAPG algorithm}\label{ssec:quiv-repr-rwapg}
                \item[$\blacksquare$] R-WAPG Intermediate form.
                \item[$\blacksquare$] R-WAPG Similar triangle form. 
                \item[$\blacksquare$] R-WAPG Momentum form. 
            Theorems: 
                \item[$\blacksquare$] R-WAPG First equivalent form. 
                \item[$\blacksquare$] R-WAPG Second equivalent form.
                \item[$\blacksquare$] R-WAPG Third equivalent form.
            Lemmas
            Todo: 
                \item Explain what is what. 
        Definitions \ref{def:r-wapg-intermediate}, \ref{def:r-wapg-st-form} and \ref{def:r-wapg-momentum-form} are three equivalent representations of the R-WAPG algorithms. 
        Propositions \ref{prop:wapg-first-equivalent-repr}, \ref{prop:wagp-st-form} and \ref{prop:r-wapg-momentum-repr} states the equivalences between the forms and the sufficient conditions for initial conditions of $x_1, v_1$ such equivalence holds. 
        \todo{$\square$ Add and explain this part. }{\hl{
        The remarks of the definitions identifies specific instances in the literatures where the Accelerated Proximal Gradient method were presented under this specific form. 
        }}

@@

            This form of APG is rarely identified in the literatures. 
            The closest algorithm that fits the form but with $\mu = 0$ is Chapter 12 of 
            \todo{$\blacksquare$ Add them \\ $\square$ Cite them. }
            {\hl{in Ryu and Yin's Book} \cite{ryu_large-scale_2022}}, 
            right after Theorem 17. 
            We created this form which makes the math that follows simpler. 
            The inspiration of using this as an intermediate representation was inspired by solving Exercise 12.1 in the same Ryu and Yin's Book. 

@@

            The word similar triangle form can be traced back to several literatures. 
            The term ``Method of Similar Triangle" was used for Algorithm (6.1.19) in Nesterov's book \cite{nesterov_lectures_2018}, but without the necessary graphical illustrations to clarify it. 
            Finally, a similar triangle for formulation of FISTA can be found in Equation (2), (3), (4) in 
            \todo{$\square$ Cite Chambolle Dossal 2015 again. }
            \cite{chambolle_convergence_2015}. 
            To see graphical visualization on why such term is used to describe the APG algorithm in the literatures, see 
            \todo{
                $\blacksquare$ Add these. $\square$ Cite these. 
            {(3.1, 4.1) in Lee et al. \cite{lee_geometric_2021} and Ahn and Sra \cite{ahn_understanding_2022}}. 
                The term ``Method of Similar Triangle'' specific to this particular representation of AGM. 
                For more information, see:  
                    \item 3.1, 4.1 in Jongmin Lee et al., ``A geometric structure of Acceleration and its role in making gradients small''
                    \item The entirety of Ahn and Sra's paper ``PPM interpretation of Nesterov's accelerated gradient. ''. 

@@

            This format fits with (2.2.19) in Nesterov's book \cite{nesterov_lectures_2018}, however, the sequence $(\alpha_k)_{k \ge 0}$ would be given by a different rule. 
            See Theorem \ref{thm:r-wapg-on-cham-doss} and Lemma \ref{lemma:inverted-fista-seq} to see a specific choice of $(\alpha_k)_{k \ge0}, (\rho_k)_{ k\ge 0}$ such this equivalent form of R-WAPG is in fact two possible variants of the FISTA algorithm.
